---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS Litters Data

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
litters_df
```

## `select` 

Choose some columns and not others. 

```{r}
select(litters_df, group, gd0_weight)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

Removing some columns with minus sign

```{r}
select(litters_df, -litter_number)
```

Renaming columns, new names go first 

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Can also use the rename function: 

```{r}
rename(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

Select helpers

```{r}
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, everything())
```

```{r}
relocate(litters_df, litter_number)
```

Select operates on columns

## `filter`

Operates on rows. Must give a condition for the rows to be kept 

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```

```{r}
filter(litters_df, gd_of_birth == 20 )
```
 Using two equal signs because we want to see if they are equal to each other and not assigning 

```{r}
filter(litters_df, !(gd_of_birth == 20))
```
Shows all those that do not equal twenty 

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```
For looking for subset where two conditions are two 

```{r}
filter(litters_df, group == "Mod8")
```

```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```
If observations are in two groups

## `mutate`

Can modify existing variable or make a new variable

```{r}
mutate(
        litters_df, 
        wt_gain = gd18_weight - gd0_weight,
        group = str_to_lower(group))
```
Changed group to lower case with str_to_lower

## `arrange`

Puts things in order. 

```{r}
arrange(litters_df, pups_born_alive)
```

```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```
Ordering by both weight and pups alive 

## `%>%` 

Piping! First a bad example with intermediate objects
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```

USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
        read_csv("./data/FAS_litters.csv") %>% 
        janitor::clean_names() %>% 
        select(-pups_survive) %>% 
        mutate(wt_gain = gd18_weight - gd0_weight) %>% 
        drop_na(gd0_weight)
```

use shift control M to insert pipe operator












