---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset. 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

The two colons after janitor tells R to pull from the Janitor package. Helps for when we want to pull just one item from package


## Take a look at the data.

Printing in the console to see everything going in.
```{r}
litters_df
```

dbl is double which is essentially a number

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```

Can also use the view function directly in the console

## Options to read csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

A good way to skip first rows or to make the first rows not the column names. 

```{r include = FALSE}
litters_df = read_csv("./data/FAS_litters.csv")
```

Option to omit missing data
 
Check out `?read_csv()` for more information. 
 
## Other file formats

Read in an excel file 
 
```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

There are some times when loading the whole library is needed instead of ::. If there are multiple sheets you have to specify which. Can specify range of import by columns and rows. 
 
 Read in a SAS file
 
```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```
 
## Comparison with Base R

What about `read.csv` ....?

Never use read.cvs, always use read_csv

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")

litters_base
litters_readr
```
 Easier to manage read_csv 
