---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data

Read in the litters dataset. 

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

The two colons after janitor tells R to pull from the Janitor package. Helps for when we want to pull just one item from package


## Take a look at the data.

Printing in the console to see everything going in.
```{r}
litters_df
```

dbl is double which is essentially a number

```{r}
head(litters_df)
```

```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```

Can also use the view function directly in the console

## Options to read csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

A good way to skip first rows or to make the first rows not the column names. 

```{r}
litters_df = read_csv("./data/FAS_litters.csv", no = c("", "NA", ".", 999))
```
 Option to omit missing data
 
 Check out `?read_csv()` for more information 